\acknowledgements{My Ph.D. journey would not have begun without the support of parents.  They are my first and best teachers in my life, who have demonstrated the courage to face challenges, perseverance through hard times, and pursuit of excellence.  This dissertation is dedicated to my parents and family for their belief in me. 

I have been very fortunate to join Law Combustion Group and work closely with Professor Chung K. Law.  I am always amazed by his dedication to work and the breadth and depth of his research.  He is a great scholar who taught me to appreciate the beauty of science, conduct rigorous research, and take initiative.  He has brought out the best in me through daily discussions about my research progress, introducing internal and external research resources to me, and exposing me to networking opportunities.  To help me develop a computational skillset, he recommended that I join the Computational Turbulent Reacting Flow Laboratory and be co-advised by Professor Michael E. Mueller since my fourth year in graduate school.  I am very grateful for his thoughtful suggestions on my career development.

I truly appreciate the patience and guidance that Professor Michael E. Mueller has given to me along my endeavor.  He spent a tremendous amount of time passing on his knowledge and skills to me and holding my hand to get me started in computational research as soon as possible.  The freedom and trust he has given to me allows me to consolidate what I have learnt in both groups, master new tools, and propel my research in new directions.  I am also very thankful for him generously sharing his experience with me on pursuing a job in academia and establishing a research group, which will benefit my future career. 

It has been a privilege to pursue my Ph.D. degree in such an intellectually challenging and nourishing environment, the Department of Mechanical and Aerospace Engineering at Princeton University.  I would like to thank the faculty, staff, and my peers in the department for their guidance, support, and companionship throughout the years.  Specifically, I am very grateful to Professors Harvey Lam and Alexander Smits for helping me prepare for my general examinations, and Professors Daniel Nosenchuck, Yiguang Ju, Craig Arnold, and Frederick Dryer for serving as my general examiners.  Furthermore, I would like to thank my Ph.D. committee members, Professors Yiguang Ju and Marcus Hultmark for their advice over the years.  I am particularly thankful to Professors Suk Ho Chung and Yiguang Ju for their feedback and suggestions to improve this dissertation.  

My colleagues in LCG and CTRFL have helped me transition from a little sister who seeks for support and suggestions to a big sister who can finally give back to the group.  I would not have mastered the counterflow experiments so quickly without the assistance of Dr. Wei Liu, Mr. Jeremy Koch, and Professor Peng Zhao.  They have not only passed on their knowledge and experience to me but also made themselves available through phone calls whenever I need their opinions.  Mr. Delin Zhu, the lab manager of LCG, more generally known as the lab magician, allows me to quickly try out new ideas and always amazes me with his masterpiece designs.  I have also enjoyed collaborating with Professor Dong Han, Mr. Wenkai Liang, and Dr. Fujia Wu on chemical kinetics.  I am indebted to my colleagues, who keep me company while doing experiments at night, challenge me during group meetings, proofread my manuscripts, help me debug programs, give me feedback during presentation rehearsals, share thoughts about research and life with me, and engage me in group activities.  I have had to walk through this Ph.D. journey by myself, but I know that I am not all by myself.       

I have been very fortunate to have the opportunity to collaborate with many outstanding researchers outside Princeton University during my Ph.D. study:  Professor Qing Chan at the University of New South Wales, Dr. Nader Qamar at FCT-Combustion, Professors Bassam Dally, Zeyad Alwahabi, and Graham Nathan at the University of Adelaide, and Professor Zhen Huang at Shanghai Jiao Tong University.   

I would also like to acknowledge several faculty and friends for the helpful discussions and technical support during my first research project on nonpremixed cool flames.  The discussions with Professor Richard Miles in MAE, Professor Claire Gmachl in EE, and Professor Marshall Long at Yale University were very helpful for the design of the detection system.  I am very grateful to Dr. Ben Wu in EE for lending me optical detectors, filters, and cables, Dr. Yu Song in EE for helping me implement the infrared camera, and Dr. Jun Xiong in Physics for helping me with the LABVIEW program for automatic control.

Finally, this thesis would not be possible without the financial support of Combustion Energy Frontier Research Center, funded by the U.S. Department of Energy, and Air Force Office of Scientific Research.  I would also like to acknowledge additional support from Princeton Climate and Energy Scholarship and Gordon Wu Prize for Excellence.  This dissertation carries T-3327 in the records of the Department of Mechanical and Aerospace Engineering.}
